Estimating Planetary Habitability via Particle Swarm Optimization of CES Production Functions.

### 2018-01-17 docs/report/draft.tex
Tables are now autogenerated and loaded.

\documentclass[9pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}

@@

\graphicspath{{figs/}}
\sisetup{output-exponent-marker=\ensuremath{\mathrm{e}}}

@@

% Useful for initial drafts.
\newenvironment{pointers}{%
  \noindent Should include,
  \begin{itemize}
    \setlength{\itemsep}{-1pt}}{%
\end{itemize}}

@@

The data set used for estimating the Habitability Scores of exoplanets, described in Section~\ref{sec:habscore}, was the
Confirmed Exoplanets Catalog maintained by the Planetary Habitability Laboratory (PHL). The catalog records observed and
modeled parameters for exoplanets confirmed by the Extrasolar Planets Encyclopedia. Table~\ref{tab:param} describes the
parameters from the PHL Exoplanets Catalog (PHL-EC) used for the experiment. Since surface temperature and eccentricity
are not recorded in Earth Units, we normalized these values by dividing them with Earth's surface temperature (288K) and
eccentricity (0.017). PHL-EC assumes an Eccentricity of 0 when unavailable. The PHL-EC records empty values for planets
whose surface temperature is not known. We chose to drop these records from the experiment.

@@

@@

\begin{figure}
  \centering
  \begin{subfigure}[b]{0.4\textwidth}
    \includegraphics[width=\textwidth]{dcdhscrs.png}
    \caption{CRS Score Distribution}\label{fig:distcdcrs}
  \end{subfigure}
  \quad
  \begin{subfigure}[b]{0.4\textwidth}
    \includegraphics[width=\textwidth]{dcdhsdrs.png}
    \caption{DRS Score Distribution}\label{fig:distcddrs}
  \end{subfigure}

  \begin{subfigure}[b]{0.4\textwidth}
    \includegraphics[width=\textwidth]{icdhscrs.png}
    \caption{CRS Iterations Distribution}\label{fig:itercdcrs}
  \end{subfigure}
  \quad
  \begin{subfigure}[b]{0.4\textwidth}
    \includegraphics[width=\textwidth]{icdhsdrs.png}
    \caption{DRS Iterations Distribution}\label{fig:itercddrs}
  \end{subfigure}
  \caption{Plots for the Cobb-Douglas Habitability Score.}\label{fig:cdhs}
\end{figure}

\begin{figure}
  \centering
  \begin{subfigure}[b]{0.4\textwidth}
    \includegraphics[width=\textwidth]{dceesacrs.png}
    \caption{CRS Score Distribution}\label{fig:distcecrs}
  \end{subfigure}
  \quad
  \begin{subfigure}[b]{0.4\textwidth}
    \includegraphics[width=\textwidth]{dceesadrs.png}
    \caption{DRS Score Distribution}\label{fig:distcedrs}
  \end{subfigure}

  \begin{subfigure}[b]{0.4\textwidth}
    \includegraphics[width=\textwidth]{iceesacrs.png}
    \caption{CRS Iterations Distribution}\label{fig:itercecrs}
  \end{subfigure}
  \quad
  \begin{subfigure}[b]{0.4\textwidth}
    \includegraphics[width=\textwidth]{iceesadrs.png}
    \caption{DRS Iterations Distribution}\label{fig:itercedrs}
  \end{subfigure}
  \caption{Plots for the Constant Elasticity Earth Similarity Approach.}\label{fig:ceesa}
\end{figure}

\begin{table}
  \centering
  \begin{subtable}{\textwidth}
    \centering
    \begin{tabular}{l r r r r r r r r r r}
      \toprule
      Name & Class & $\alpha$ & $\beta$ & $Y_i$ & $i_i$ & $\gamma$ & $\delta$ & $Y_s$ & $i_s$ & $\mathit{CDHS}$\\
      \midrule
      GJ 176 b & non & 0.460 & 0.540 & 1.90 & 50 & 0.107 & 0.893 & 2.11 & 61 & 1.90\\
      GJ 667 C b & non & 0.423 & 0.577 & 1.71 & 58 & 0.692 & 0.308 & 1.81 & 54 & 1.71\\
      GJ 667 C e & psy & 0.129 & 0.871 & 1.40 & 50 & 0.258 & 0.742 & 1.39 & 55 & 1.40\\
      GJ 667 C f & psy & 0.534 & 0.466 & 1.40 & 48 & 0.865 & 0.135 & 1.39 & 47 & 1.40\\
      GJ 3634 b & non & 0.409 & 0.591 & 1.89 & 58 & 0.724 & 0.276 & 2.09 & 48 & 1.89\\
      HD 20794 c & non & 0.260 & 0.740 & 1.35 & 50 & 0.096 & 0.904 & 1.34 & 58 & 1.35\\
      HD 40307 e & non & 0.168 & 0.832 & 1.50 & 49 & 0.636 & 0.364 & 1.53 & 63 & 1.50\\
      HD 40307 f & non & 0.702 & 0.298 & 1.52 & 68 & 0.303 & 0.697 & 1.55 & 45 & 1.52\\
      HD 40307 g & psy & 0.964 & 0.036 & 1.82 & 51 & 0.083 & 0.917 & 1.98 & 55 & 1.82\\
      Kepler-186 f & hyp & 0.338 & 0.662 & 1.17 & 50 & 0.979 & 0.021 & 1.12 & 40 & 1.17\\
      Proxima Cen b & psy & 0.515 & 0.484 & 1.12 & 37 & 0.755 & 0.245 & 1.07 & 00 & 1.12\\
      TRAPPIST-1 b & non & 0.319 & 0.681 & 1.09 & 00 & 0.801 & 0.199 & 0.89 & 00 & 1.09\\
      TRAPPIST-1 c & non & 0.465 & 0.535 & 1.06 & 00 & 0.935 & 0.065 & 1.14 & 26 & 1.06\\
      TRAPPIST-1 d & mes & 0.635 & 0.365 & 0.77 & 34 & 0.475 & 0.525 & 0.73 & 47 & 0.77\\
      TRAPPIST-1 e & psy & 0.145 & 0.855 & 0.92 & 00 & 0.897 & 0.103 & 0.83 & 55 & 0.92\\
      TRAPPIST-1 g & hyp & 0.226 & 0.774 & 1.13 & 43 & 0.876 & 0.124 & 1.09 & 00 & 1.13\\
      \bottomrule
    \end{tabular}
    \caption{Estimated habitability scores under CRS.}\label{tab:cdhscrs}
  \end{subtable}

@@

  \begin{subtable}{\textwidth}
    \centering
    \begin{tabular}{l r r r r r r r r r r}
      \toprule
      Name & Class & $\alpha$ & $\beta$ & $Y_i$ & $i_i$ & $\gamma$ & $\delta$ & $Y_s$ & $i_s$ & $\mathit{CDHS}$\\
      \midrule
      GJ 176 b & non & 0.395 & 0.604 & 1.90 & 59 & 0.372 & 0.627 & 2.11 & 56 & 1.90\\
      GJ 667 C b & non & 0.781 & 0.218 & 1.71 & 58 & 0.902 & 0.097 & 1.81 & 57 & 1.71\\
      GJ 667 C e & psy & 0.179 & 0.820 & 1.40 & 49 & 0.234 & 0.765 & 1.39 & 60 & 1.40\\
      GJ 667 C f & psy & 0.704 & 0.295 & 1.40 & 64 & 0.398 & 0.601 & 1.39 & 61 & 1.40\\
      GJ 3634 b & non & 0.602 & 0.397 & 1.89 & 59 & 0.429 & 0.570 & 2.09 & 77 & 1.89\\
      HD 20794 c & non & 0.014 & 0.985 & 1.35 & 50 & 0.116 & 0.883 & 1.34 & 45 & 1.35\\
      HD 40307 e & non & 0.752 & 0.247 & 1.50 & 60 & 0.677 & 0.322 & 1.53 & 50 & 1.50\\
      HD 40307 f & non & 0.887 & 0.112 & 1.52 & 51 & 0.261 & 0.738 & 1.55 & 60 & 1.52\\
      HD 40307 g & psy & 0.300 & 0.699 & 1.82 & 62 & 0.785 & 0.214 & 1.98 & 56 & 1.82\\
      Kepler-186 f & hyp & 0.073 & 0.926 & 1.17 & 46 & 0.740 & 0.259 & 1.12 & 51 & 1.17\\
      Proxima Cen b & psy & 0.045 & 0.954 & 1.12 & 57 & 0.216 & 0.783 & 1.07 & 53 & 1.12\\
      TRAPPIST-1 b & non & 0.102 & 0.897 & 1.09 & 41 & 0.000 & 0.000 & 1.00 & 65 & 1.09\\
      TRAPPIST-1 c & non & 0.471 & 0.528 & 1.06 & 44 & 0.227 & 0.772 & 1.14 & 57 & 1.06\\
      TRAPPIST-1 d & mes & 0.000 & 0.000 & 1.00 & 67 & 0.000 & 0.000 & 1.00 & 59 & 1.00\\
      TRAPPIST-1 e & psy & 0.000 & 0.000 & 1.00 & 55 & 0.000 & 0.000 & 1.00 & 57 & 1.00\\
      TRAPPIST-1 g & hyp & 0.888 & 0.111 & 1.13 & 47 & 0.949 & 0.050 & 1.09 & 46 & 1.13\\
      \bottomrule
    \end{tabular}
    \caption{Estimated habitability scores under DRS.}\label{tab:cdhsdrs}
  \end{subtable}
  \caption{Cobb-Douglas Habitability Scores as estimated by \pso.\\
    \footnotesize $\alpha$, $\beta$, $\gamma$ and $\delta$ record the parameters of Equation~\ref{eq:cdhscrs} in
    Table~\ref{tab:cdhscrs} and the parameters of Equation~\ref{eq:cdhsdrs} in Table~\ref{tab:cdhsdrs}. $Y_i$ and $Y_s$
    record the maxima for the objective functions. $i_i$ and $i_s$ specify the number of iterations taken to converge to
    a stable $\gbest$ value. Under the Class column there are four categories for the planets --- Psychroplanets (psy),
    Mesoplanets (mes), Non-Habitable planets (non) and Hypopsychroplanets (hyp).  $\mathit{CDHS}$ is the final score
    with $w_i=0.99$ and $w_s=0.01$.
  }
\end{table}

\begin{table}
  \centering
  \begin{subtable}{\textwidth}
    \centering
    \begin{tabular}{l r r r r r r r r r r}
      \toprule
      Name & Class & $r$ & $d$ & $t$ & $v$ & $e$ & $\rho$ & $\eta$ & $\mathit{CEESA}$ & $i$\\
      \midrule
      GJ 176 b & non & 0.304 & 0.001 & 0.375 & 0.271 & 0.050 & 0.467 & 0.808 & 1.52 &  85\\
      GJ 667 C b & non & 0.297 & 0.010 & 0.318 & 0.052 & 0.322 & 0.682 & 0.730 & 2.36 &  90\\
      GJ 667 C e & psy & 0.230 & 0.286 & 0.137 & 0.199 & 0.148 & 0.551 & 0.906 & 1.14 &  85\\
      GJ 667 C f & psy & 0.397 & 0.035 & 0.152 & 0.402 & 0.014 & 0.793 & 0.999 & 1.31 & 100\\
      GJ 3634 b & non & 0.178 & 0.175 & 0.005 & 0.194 & 0.447 & 0.894 & 0.657 & 2.07 &  94\\
      HD 20794 c & non & 0.073 & 0.142 & 0.452 & 0.190 & 0.144 & 0.953 & 0.635 & 1.20 &  78\\
      HD 40307 e & non & 0.156 & 0.307 & 0.185 & 0.033 & 0.319 & 0.428 & 0.939 & 2.69 &  88\\
      HD 40307 f & non & 0.272 & 0.231 & 0.064 & 0.305 & 0.127 & 0.676 & 0.802 & 1.28 &  77\\
      HD 40307 g & psy & 0.113 & 0.219 & 0.066 & 0.454 & 0.148 & 0.711 & 0.991 & 3.26 &  92\\
      Kepler-186 f & hyp & 0.039 & 0.159 & 0.116 & 0.329 & 0.357 & 0.253 & 0.919 & 1.35 &  70\\
      Proxima Cen b & psy & 0.272 & 0.173 & 0.284 & 0.193 & 0.079 & 0.615 & 0.114 & 0.99 &  75\\
      TRAPPIST-1 b & non & 0.488 & 0.151 & 0.039 & 0.193 & 0.129 & 0.151 & 0.014 & 0.99 &  87\\
      TRAPPIST-1 c & non & 0.172 & 0.236 & 0.275 & 0.242 & 0.075 & 0.969 & 0.962 & 1.06 &  80\\
      TRAPPIST-1 d & mes & 0.106 & 0.308 & 0.075 & 0.218 & 0.293 & 0.844 & 0.017 & 0.99 &  93\\
      TRAPPIST-1 e & psy & 0.189 & 0.266 & 0.192 & 0.094 & 0.260 & 0.371 & 0.006 & 0.99 &  84\\
      TRAPPIST-1 g & hyp & 0.326 & 0.186 & 0.143 & 0.278 & 0.067 & 0.315 & 0.021 & 1.00 &  76\\
      \bottomrule
    \end{tabular}
    \caption{Estimated habitability scores under DRS.}\label{tab:ceesadrs}
  \end{subtable}

@@

  \begin{subtable}{\textwidth}
    \centering
    \begin{tabular}{l r r r r r r r r r}
      \toprule
      Name & Class & $r$ & $d$ & $t$ & $v$ & $e$ & $\rho$ & $\mathit{CEESA}$ & $i$\\
      \midrule
      GJ 176 b & non & 0.194 & 0.020 & 0.315 & 0.465 & 0.006 & 0.398 & 1.88 &  86\\
      GJ 667 C b & non & 0.162 & 0.289 & 0.090 & 0.087 & 0.372 & 0.836 & 3.54 & 107\\
      GJ 667 C e & psy & 0.373 & 0.032 & 0.134 & 0.304 & 0.157 & 0.217 & 1.25 &  71\\
      GJ 667 C f & psy & 0.394 & 0.006 & 0.043 & 0.360 & 0.196 & 0.490 & 1.44 &  81\\
      GJ 3634 b & non & 0.351 & 0.122 & 0.006 & 0.069 & 0.453 & 0.439 & 2.89 &  96\\
      HD 20794 c & non & 0.101 & 0.077 & 0.691 & 0.071 & 0.059 & 0.756 & 1.58 &  94\\
      HD 40307 e & non & 0.069 & 0.091 & 0.097 & 0.173 & 0.569 & 0.768 & 5.29 &  94\\
      HD 40307 f & non & 0.285 & 0.161 & 0.053 & 0.443 & 0.058 & 0.342 & 1.42 &  73\\
      HD 40307 g & psy & 0.156 & 0.010 & 0.081 & 0.302 & 0.451 & 0.612 & 7.15 &  94\\
      Kepler-186 f & hyp & 0.036 & 0.017 & 0.082 & 0.383 & 0.483 & 0.929 & 1.68 &  85\\
      Proxima Cen b & psy & 0.352 & 0.383 & 0.103 & 0.059 & 0.103 & 0.936 & 0.89 &  83\\
      TRAPPIST-1 b & non & 0.148 & 0.147 & 0.344 & 0.269 & 0.093 & 0.767 & 0.94 &  81\\
      TRAPPIST-1 c & non & 0.038 & 0.060 & 0.575 & 0.321 & 0.005 & 0.602 & 1.17 &  86\\
      TRAPPIST-1 d & mes & 0.023 & 0.065 & 0.475 & 0.391 & 0.045 & 0.830 & 0.84 &  79\\
      TRAPPIST-1 e & psy & 0.176 & 0.464 & 0.253 & 0.103 & 0.004 & 0.920 & 0.86 &  81\\
      TRAPPIST-1 g & hyp & 0.060 & 0.086 & 0.310 & 0.540 & 0.004 & 0.848 & 0.97 &  86\\
      \bottomrule
    \end{tabular}
    \caption{Estimated habitability scores under CRS.}\label{tab:ceesacrs}
  \end{subtable}
  \caption{Constant Elasticity Earth Similarity Approach scores as estimated by \pso.\\
      \footnotesize $r$, $d$, $t$, $v$, $e$, $\rho$ and $eta$ record the parameters of Equation~\ref{eq:ceesadrs} in
      Table~\ref{tab:ceesadrs} and the parameters of Equation~\ref{eq:ceesacrs} in Table~\ref{tab:ceesacrs}. However,
      since under the CRS constraint, $\eta=1$, there is no need for the parameter $\eta$ in the problem. The column
      $\mathit{CEESA}$ records the maxima of the objective function. $i$ specifies the number of iterations taken to
      converge to the maxima.
    }
\end{table}

\section{Conclusions}
\begin{pointers}
\item Why is the speed so important?
\item Parallelizable.
\end{pointers}

>>>>

\documentclass[10pt]{article}

\usepackage[margin=1.2in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}

@@

\graphicspath{{figs/}}
% \sisetup{output-exponent-marker=\ensuremath{\mathrm{e}}}

@@

@@

@@

\begin{figure}
  \centering
  \begin{subfigure}[b]{0.38\textwidth}
    \includegraphics[width=\textwidth]{dcdhscrs.png}
    \caption{CRS Score Distribution}\label{fig:distcdcrs}
  \end{subfigure}
  \quad
  \begin{subfigure}[b]{0.38\textwidth}
    \includegraphics[width=\textwidth]{dcdhsdrs.png}
    \caption{DRS Score Distribution}\label{fig:distcddrs}
  \end{subfigure}

  \begin{subfigure}[b]{0.38\textwidth}
    \includegraphics[width=\textwidth]{icdhscrs.png}
    \caption{CRS Iterations Distribution}\label{fig:itercdcrs}
  \end{subfigure}
  \quad
  \begin{subfigure}[b]{0.38\textwidth}
    \includegraphics[width=\textwidth]{icdhsdrs.png}
    \caption{DRS Iterations Distribution}\label{fig:itercddrs}
  \end{subfigure}
  \caption{Plots for the Cobb-Douglas Habitability Score.}\label{fig:cdhs}
\end{figure}

\begin{figure}
  \centering
  \begin{subfigure}[b]{0.38\textwidth}
    \includegraphics[width=\textwidth]{dceesacrs.png}
    \caption{CRS Score Distribution}\label{fig:distcecrs}
  \end{subfigure}
  \quad
  \begin{subfigure}[b]{0.38\textwidth}
    \includegraphics[width=\textwidth]{dceesadrs.png}
    \caption{DRS Score Distribution}\label{fig:distcedrs}
  \end{subfigure}

  \begin{subfigure}[b]{0.38\textwidth}
    \includegraphics[width=\textwidth]{iceesacrs.png}
    \caption{CRS Iterations Distribution}\label{fig:itercecrs}
  \end{subfigure}
  \quad
  \begin{subfigure}[b]{0.38\textwidth}
    \includegraphics[width=\textwidth]{iceesadrs.png}
    \caption{DRS Iterations Distribution}\label{fig:itercedrs}
  \end{subfigure}
  \caption{Plots for the Constant Elasticity Earth Similarity Approach.}\label{fig:ceesa}
\end{figure}

The data set used for estimating the Habitability Scores of exoplanets, described in Section~\ref{sec:habscore}, was the
Confirmed Exoplanets Catalog maintained by the Planetary Habitability Laboratory (PHL). The catalog records observed and
modeled parameters for exoplanets confirmed by the Extrasolar Planets Encyclopedia. Table~\ref{tab:param} describes the
parameters from the PHL Exoplanets Catalog (PHL-EC) used for the experiment. Since surface temperature and eccentricity
are not recorded in Earth Units, we normalized these values by dividing them with Earth's surface temperature
(\SI{288}{\kelvin}) and eccentricity (\num{0.017}). PHL-EC assumes an Eccentricity of 0 when unavailable. The PHL-EC
records empty values for planets whose surface temperature is not known. We chose to drop these records from the
experiment.

@@

As for real time taken to converge, PSO took \SI{666.85}{\second} ($\approx\SI{11}{\minute}\ \SI{7}{\second}$) to
estimate the CDHS under CRS for \num{1683} exoplanets, at an average of \SI{198.11}{\milli\second} for each planet for
each part of the CDHS.\@ For CDHS under DRS, it took \SI{638.69}{\second} ($\approx\SI{10}{\minute}\ \SI{39}{\second}$)
at an average of \SI{189.75}{\milli\second} for each part of the CDHS.\@ The CEESA estimates, requiring one calculation,
took a little over half the total time to execute. Under DRS it took a total of \SI{370.86}{\second}
($\approx\SI{6}{\minute}\ \SI{11}{\second}$) at \SI{220.36}{\milli\second} per planet, while under CRS it took
\SI{356.92}{\second} ($\approx\SI{5}{\minute}\ \SI{57}{\second}$) at \SI{212.07}{\milli\second} per planet. This is
considerably fast considering the number of parameters and constraints for the problem.

\begin{table}
  \centering
  \begin{subtable}{\textwidth}
    \centering
    \include{./tabs/cdhscrs}
    \caption{Estimated habitability scores under CRS.}\label{tab:cdhscrs}
  \end{subtable}

@@

  \begin{subtable}{\textwidth}
    \centering
    \include{./tabs/cdhsdrs}
    \caption{Estimated habitability scores under DRS.}\label{tab:cdhsdrs}
  \end{subtable}
  \caption{Cobb-Douglas Habitability Scores as estimated by \pso.\\
    \footnotesize $\alpha$, $\beta$, $\gamma$ and $\delta$ record the parameters of Equation~\ref{eq:cdhscrs} in
    Table~\ref{tab:cdhscrs} and the parameters of Equation~\ref{eq:cdhsdrs} in Table~\ref{tab:cdhsdrs}. $Y_i$ and $Y_s$
    record the maxima for the objective functions. $i_i$ and $i_s$ specify the number of iterations taken to converge to
    a stable $\gbest$ value. Under the Class column there are four categories for the planets --- Psychroplanets (psy),
    Mesoplanets (mes), Non-Habitable planets (non) and Hypopsychroplanets (hyp).  $\mathit{CDHS}$ is the final score
    with $w_i=0.99$ and $w_s=0.01$.
  }
\end{table}

\begin{table}
  \centering
  \begin{subtable}{\textwidth}
    \centering
    \include{./tabs/ceesadrs}
    \caption{Estimated habitability scores under DRS.}\label{tab:ceesadrs}
  \end{subtable}

@@

  \begin{subtable}{\textwidth}
    \centering
    \include{./tabs/ceesacrs}
    \caption{Estimated habitability scores under CRS.}\label{tab:ceesacrs}
  \end{subtable}
  \caption{Constant Elasticity Earth Similarity Approach scores as estimated by \pso.\\
      \footnotesize $r$, $d$, $t$, $v$, $e$, $\rho$ and $eta$ record the parameters of Equation~\ref{eq:ceesadrs} in
      Table~\ref{tab:ceesadrs} and the parameters of Equation~\ref{eq:ceesacrs} in Table~\ref{tab:ceesacrs}. However,
      since under the CRS constraint, $\eta=1$, there is no need for the parameter $\eta$ in the problem. The column
      $\mathit{CEESA}$ records the maxima of the objective function. $i$ specifies the number of iterations taken to
      converge to the maxima.
    }
\end{table}

\section{Conclusions}
% - Why is the speed so important?
% - Parallelizable.
